\documentclass[11pt]{article}

\newcommand{\yourname}{}
\newcommand{\yourcollaborators}{}

\def\comments{0}

%format and packages

%\usepackage{algorithm, algorithmic}

\usepackage{ulem}
\usepackage{epsfig, graphicx}
\usepackage[noend]{algpseudocode}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{framed}
\usepackage{verbatim}
\usepackage[margin=1.1in]{geometry}
\usepackage{microtype}
\usepackage{kpfonts}
\usepackage{palatino}
	\DeclareMathAlphabet{\mathtt}{OT1}{cmtt}{m}{n}
	\SetMathAlphabet{\mathtt}{bold}{OT1}{cmtt}{bx}{n}
	\DeclareMathAlphabet{\mathsf}{OT1}{cmss}{m}{n}
	\SetMathAlphabet{\mathsf}{bold}{OT1}{cmss}{bx}{n}
	\renewcommand*\ttdefault{cmtt}
	\renewcommand*\sfdefault{cmss}
	\renewcommand{\baselinestretch}{1.05}
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{DarkGreen}{rgb}{0.15,0.5,0.15}
\definecolor{DarkRed}{rgb}{0.6,0.2,0.2}
\definecolor{DarkBlue}{rgb}{0.2,0.2,0.6}
\definecolor{DarkPurple}{rgb}{0.4,0.2,0.4}
\usepackage[pdftex]{hyperref}
\hypersetup{
	linktocpage=true,
	colorlinks=true,				% false: boxed links; true: colored links
	linkcolor=DarkBlue,		% color of internal links
	citecolor=DarkBlue,	% color of links to bibliography
	urlcolor=DarkBlue,		% color of external links
}

\usepackage[boxruled,vlined,nofillcomment]{algorithm2e}
	\SetKwProg{Fn}{Function}{\string:}{}
	\SetKwFor{While}{While}{}{}
	\SetKwFor{For}{For}{}{}
	\SetKwIF{If}{ElseIf}{Else}{If}{:}{ElseIf}{Else}{:}
	\SetKw{Return}{Return}
	

%enclosure macros
\newcommand{\paren}[1]{\ensuremath{\left( {#1} \right)}}
\newcommand{\bracket}[1]{\ensuremath{\left\{ {#1} \right\}}}
\renewcommand{\sb}[1]{\ensuremath{\left[ {#1} \right\]}}
\newcommand{\ab}[1]{\ensuremath{\left\langle {#1} \right\rangle}}

%probability macros
\newcommand{\ex}[2]{{\ifx&#1& \mathbb{E} \else \underset{#1}{\mathbb{E}} \fi \left[#2\right]}}
\newcommand{\pr}[2]{{\ifx&#1& \mathbb{P} \else \underset{#1}{\mathbb{P}} \fi \left[#2\right]}}
\newcommand{\var}[2]{{\ifx&#1& \mathrm{Var} \else \underset{#1}{\mathrm{Var}} \fi \left[#2\right]}}

%useful CS macros
\newcommand{\poly}{\mathrm{poly}}
\newcommand{\polylog}{\mathrm{polylog}}
\newcommand{\zo}{\{0,1\}}
\newcommand{\pmo}{\{\pm1\}}
\newcommand{\getsr}{\gets_{\mbox{\tiny R}}}
\newcommand{\card}[1]{\left| #1 \right|}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\negl}{\mathrm{negl}}
\newcommand{\eps}{\varepsilon}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\eqand}{\qquad \textrm{and} \qquad}
\newcommand{\ind}[1]{\mathbb{I}\{#1\}}
\newcommand{\sslash}{\ensuremath{\mathbin{/\mkern-3mu/}}}

%info theory macros
\newcommand{\SD}{\mathit{SD}}
\newcommand{\sd}[2]{\SD\left( #1 , #2 \right)}
\newcommand{\KL}{\mathit{KL}}
\newcommand{\kl}[2]{\KL\left(#1 \| #2 \right)}
\newcommand{\CS}{\ensuremath{\chi^2}}
\newcommand{\cs}[2]{\CS\left(#1 \| #2 \right)}
\newcommand{\MI}{\mathit{I}}
\newcommand{\mi}[2]{\MI\left(~#1~;~#2~\right)}

%mathbb
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
%mathcal
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cO}{\mathcal{O}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\cQ}{\mathcal{Q}}
\newcommand{\cR}{\mathcal{R}}
\newcommand{\cS}{\mathcal{S}}
\newcommand{\cU}{\mathcal{U}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cZ}{\mathcal{Z}}

\newcommand{\hs}{\hspace{0.2in}}
%theorem macros
\newtheorem{thm}{Theorem}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{fact}[thm]{Fact}
\newtheorem{clm}[thm]{Claim}
\newtheorem{rem}[thm]{Remark}
\newtheorem{coro}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{conj}[thm]{Conjecture}
	\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}

\theoremstyle{theorem}
\newtheorem{prob}{Problem}


\newcommand{\course}{CS 3000: Algorithms \& Data}
\newcommand{\semester}{Spring 2024}

\newcommand{\hwnum}{2}
\newcommand{\hwdue}{Saturday, February 3 at 11:59pm via Gradescope}

\definecolor{cit}{rgb}{0.05,0.2,0.45} 

\newif\ifsolution

\solutiontrue

\ifsolution
\newcommand{\solution}[1]{\medskip\noindent{\color{DarkBlue}\textbf{Solution:}} #1}
\else
\newcommand{\solution}[1]{}
\fi

\begin{document}
{\Large 
\begin{center} \course\ --- \semester\ \end{center}}
{\large
\vspace{10pt}
\noindent Homework~\hwnum \vspace{2pt}\\
Due~\hwdue}

\vspace{15pt}
\bigskip
{\large
\noindent Name: \yourname \vspace{2pt}\\ Collaborators: \yourcollaborators}

\vspace{15pt}
\begin{itemize}

\item
  Make sure to put your name on the first page.  If you are using the
  \LaTeX~template we provided, then you can make sure it appears by
  filling in the \texttt{yourname} command.

\item This homework is due~\hwdue.  No late assignments will be accepted.  Make sure to submit something before the deadline.

\item Solutions must be typeset.  If you need to draw any diagrams,
  you may draw them by hand as long as they are embedded in the PDF.
  We recommend that you use \LaTeX, in which case it would be best to
  use the source file for this assignment to get started.

\item We encourage you to work with your classmates on the homework
  problems, but also urge you to attempt all of the problems by
  yourself first. \emph{If you do collaborate, you must write all
    solutions by yourself, in your own words.}  Do not submit anything
  you cannot explain.  Please list all your collaborators in your
  solution for each problem by filling in the
  \texttt{yourcollaborators} command.

\item Finding solutions to homework problems on the web, or by asking
  students not enrolled in the class is strictly forbidden.

\end{itemize}
\newpage



%%%%%  PROBLEM 1
\newpage
\begin{prob} (10 points) Karatsuba Example \end{prob}

\noindent Carry out Karatsuba's Algorithm to compute $47 \cdot 63$.  What are the inputs for each recursive call, what does that recursive call return, and how do we compute the final product?

\solution



%%%%%% PROBLEM 2
\newpage
\begin{prob} (13 points) Recursion Tree \end{prob}

\noindent Consider the following recurrence:

$$T(n) = 3T(n/3) + 3T(n/2) + n^2$$
$$T(1) = C$$

We will show that $T(n) = O(n^{\log_2 (13/3)})$. To do this, start by examining the first three levels of the recursion tree, showing how to compute the amount of work at each level. From here, establish a formula for the amount of work on level $i$. Then, determine the last level of the recursion tree (note that it is sufficient to focus on the largest piece at level $i$, as we are only concerned with a Big-O bound). Finally, construct a summation which totals the amount of work over all levels and show why this summation is $T(n) = O(n^{\log_2 (13/3)})$.

You are welcome to embed a photo of a hand draw image into your LaTeX file\footnote{\textbackslash includegraphics is useful for this}.\\

\solution

\newpage

%%%%%% PROBLEM 3
\begin{prob}
  \label{prob:sox}
(14 + 5 = 19 points)  Help thy neighbor
\end{prob}

\noindent  2023 was a tough year for the Red Sox. Looking to find some silver linings in the difficult season, manager Alex Cora has asked you to help identify the best stretch of the season for the team. To do this, the analytics department has shared with you the run differentials from $n$ games this season. For each game, this value can be positive (if the Sox won) or negative (if they lost). For example, if they won 10-4, then lost 6-2, then won 3-2, the run differentials would be 6, -4, 1. Alex would like you to find the stretch of consecutive games such that the total run differential is maximized. The total run differential of a stretch of games is simply the sum of the run differentials of those games. 

For example, consider the following:\\

$n=7$,
differentials: $-1, 5, -2, 1, 4, -3, 1$

Here, the optimal solution is from game 2 to game 5, and the total run differential is 8.





\begin{enumerate}[label=(\alph*)]
\item Write (in pseudocode) a divide and conquer algorithm which solves this problem in $O(n \log n)$ time. Provide a few sentences describing your approach. Note that your algorithm should output the range of games that yields the maximum total run differential.

\solution




\item Write a recurrence for the runtime of your algorithm and solve it. You may use any method for solving the recurrence that we have discussed in class. Justify your recurrence.



\solution 

\end{enumerate}

\newpage


%%%%%  PROBLEM 4
\begin{prob}
  \label{prob:equal}
(2 + 2 + 10 + 5 = 19 points) Lucky coincidence
\end{prob}

Given a sorted array of distinct integers $A[1,\ldots,n]$, you want to find out whether there is an index $i$ such that $A[i] = i$. For example, in the array $A=\{3,5,6\}$, there is no $i$ such that $A[i]=i$ and the algorithm should return ``no index found''. On the other hand, for the array $A={0, 2, 7}$ we have $A[2]=2$ and the algorithm should return $i=2$.
\begin{enumerate}[label=(\alph*)]
\item What is the answer if $A[1] > 1$?

\solution{
  
}

\item What is the answer if $A[n] < n$?

\solution{
  
}

\item Give an efficient divide-and-conquer algorithm for this problem. Hint: check if $A[\lfloor n/2\rfloor]=\lfloor n/2\rfloor$. If not, try to recurse on a smaller problem.

\solution{

}

\item Write the recurrence of the running time and solve it using the master theorem.

\solution{
}
\end{enumerate}

\newpage

%%%%%  PROBLEM 5
\begin{prob}
  \label{prob:potatoes}
(14 + 5 = 19 points) Deep in tot
\end{prob}


\noindent You have been put in charge of judging the potato sorting contest at this year's Eastern Idaho State Fair. Each contestant is tasked with sorting $n$ potatoes by weight in ascending order, without the assistance of any equipment. The judging is as follows:

\begin{itemize}
\item Given an ordering of $n$ potatoes, a \emph{mistake} consists of two potatoes $i$ and $j$ such that potato $i$ is before potato $j$ in ther order, but $i$ weighs more than $j$.
\end{itemize}

You need to be able to quickly calculate the number of mistakes in a given ordering. The runtime of your algorithm should be $O(n \log(n))$ for full credit. You have the ability to compare the weights of two potatoes in $O(1)$ time.



\begin{enumerate}[label=(\alph*)]
\item Describe a divide-and-conquer algorithm (in pseudocode) which returns the number of mistakes. Provide a complete written description of your approach.

\solution



\item State the worst-case asymptotic running time of your approach.  Write a recurrence which captures the worst-case runtime of your algorithm and solve it using any method that we've seen in class.

\solution 


\end{enumerate}


\end{document}
